The Three-HITs Theorem\footnote{This work was partially supported by COST Action CA15123.}

### 2017-03-21 Paper/Paper.tex
Finished elimination rule and computation rule

\newcommand{\frec}[0]{\!\operatorname{rec}}
\newcommand{\rec}[0]{\operatorname{rec}}
\newcommand{\nonrec}[0]{\operatorname{nonrec}}

@@

Now we will construct the map $A_{\rec} \> (\hocolim \> F \> f) \rightarrow \hocolim \> F \> f$, and by Lemma \ref{lem:colim_commute_poly} it suffices to make a map $\hocolim \> (A_{\rec} \comp F) \> (A_{\rec} \> f) \rightarrow \hocolim \> F \> f$.
For this we use the recursion rule of $\hocolim$, and we start with the following string of maps
\[
\xymatrix
{
		A_{\rec}(F \> n) \ar[r]^-{c_{\rec}'} 
		& H_{\rec}(F \> n) \ar[rr]^{\term \comp \inn_2} &
		& H_{\Con}(F \> n) \ar[r]^-{\inC}
		& H_{\Approx}(F \> n) = F(n + 1)
}
\]
Composing this map with $\inc$, gives maps $A_{\rec}(F \> n) \rightarrow \hocolim \> F \> f$ for all $n : \mathbb{N}$.

@@

To finish the proof, we need to show that $\com_*(\ap \> (\inc \> n) \> (\glueC \> x)) = \ap \> (\inc \> (n+1)) \> (\glueC \> ((B_i \>(f \>n)) \> x))$.
This follows from the following computation.
\begin{equation*}
\begin{split}
\ap &\> (\inc \> (n+1)) \> (\glueC \> ((B_i \>(f \> n)) \> x))\\
&= \ap \> (\inc \> (n+1)) \> (\ap \> (f \> n) \>\glueC \> x))\\
&= \ap \> (\inc \> (n+1) \circ (f \> n)) \>\glueC \> x))\\
&=\com_* \> (\ap \> (\inc \> n) \> (\glueC \> x))
\end{split}
\end{equation*}
The first step follows from the third added coherency, the second from Lemma 2.2.2 in \cite{hottbook}, and the last step from Lemma \ref{lem:pathext}.

\subsection{Elimination Rule}
For the next step we define the right eliminator for $\hocolim \> F \> f$, so suppose that we are given
\[
Y : \hocolim \> F \> f \rightarrow \Type,
\]
\[
c_{Y, \nonrec} : \prod (a : A_{\nonrec}), Y(c_{\nonrec} \> a),
\]
\[
c_{Y, \rec} : \prod (x : A_{\rec} \> \hocolim \> F \> f), \overline{A_{\rec}} \> Y \> x \rightarrow Y(c_{\rec} \> x),
\]
\[
q_{Y, i} : \prod (x : B_i \> H)(h_x : \overline{B_i} \> Y \> x), \hatt_i =_{p_j \, x} \hatr_j.
\]
In order to make a map $h : \prod (x : \hocolim \> F \> f), Y \> x$, we use the induction principle of $\hocolim \> F \> f$, and for that we first need to make maps $h \> n : \prod (x : F \> n), Y(\inc \> n \> x)$ for $n : \mathbb{N}$.

@@

Let us do this in the general case.
Suppose, we have a map $g : P \rightarrow \hocolim \> F \> f$, and that we already constructed $\varphi : \prod (x : P), Y(g \> x)$.
Our goal is to extend the map into $\varphi' : \prod (x : H_{\rec} P), Y(c_{Y, \rec} \> x)$, and for that we use $H_{\rec}$-induction.
Note that for each $c'_{\rec} \> x$ with $x : A_{\rec} \> P$ we have the type $Y(c_{\rec} \> (A_{\rec} \> g \> x))$, and thus we have a type family on $H_{\rec}$.
Now let $x : A_{\rec} \> P$ and $y : \overline{A_{\rec}} \> Y \> x$ be given.
Then we need to give an element of the type $Y(c_{\rec} \> (A_{\rec} \> g \> x))$ for which we take $c_{Y,\rec} \> (A_{\rec} \> g \> x) \> y$.

@@

Next we check the coherency conditions, and we start with the first.
We need to check that $\inn_1(\inC \> x)$ and $H_{\Con} \> \inC \> x$ get mapped to the same element for $x : H_{\Con}(F \> n)$ by $h_2$.
To do so, we use a case distinction on $H_{\Con}(F \> n)$, and for the case $x \equiv \inn_1 \> y$ with $y : F \> n$, we have the following equations 
\begin{equation*}
\begin{split}
h_2 \> (\inC(\inn_1(\inC \> (\inn_1 \> y))))
&\equiv
h_1 \> (\inn_1(\inC \> (\inn_1 \> y))) \\
&\equiv 
h_1 \> (\inn_1(f \> n \> y))\\
&\equiv 
h_1 \> (\inn_1(\inC(\inn_1 \> y))),
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
h_2 \> (\inC(H_{\Con} \> (\inC \circ \inn_1) \> (\inn_1 \> y)))
&\equiv
h_1 \> (H_{\Con} \> (\inC \circ \inn_1) \> (\inn_1 \> y))\\
&\equiv 
h_1 \> (\inn_1(\inC(\inn_1 \> y)))
\end{split}
\end{equation*}
Here we use that $f \> n \> x$ is defined to be $\inC(\inn_1 \> x)$.
Since they are definitionally equal, we can just use the path $\refl$.

@@

To finish the proof, we need to give the image for the path $\com$.
More concretely, we need to show that $\inc \> n \> x$ and $\inc \> (n+1) \> (f \> x)$ are mapped to the same element.
The map $f : F \> n \rightarrow F(n + 1)$ is defined by the pushout as follows
\[
\xymatrix
{
	(\two \times B_i \> (F(n+1))) + H_{\Con} \>  (F \> n) \ar[d]_{\overline{t_i} + \overline{r_i} + H_{\Con} \> f} \ar[rr]^-{\delta + \inl} & & (I^1 \times B_i \> (F(n+1))) + F(n+1)  \ar[d]^{\inr}\\
	H_{\Con} \> (F(n+1)) \ar[rr]_{\inl} & & F(n+2)
}
\]
Since the map to $Y$ was defined as the pushout map, it will commute automatically.

All in all, we have acquired a map $\prod (x : \hocolim \> F \> f), Y \> x$, and this way we defined the right eliminator for $\hocolim \> F \> f$.
We shall call the eliminator $H\frec$.

\subsection{Computation Rules}
Lastly, we show that this eliminator also satisfies the computation rules.
First, we prove that for each $t : A_{\rec}(\hocolim \> F \> f)$ that $H\frec(c_{\rec} \> t) \equiv f_i \> t \> (\overline{A_{\rec}} \> H\frec \> t)$.
Again we use that colimits commute with polynomials.

Let $n : \mathbb{N}$ and $x : A_{\rec}(F \> n)$.
Then by the computation rules we have
\begin{equation*}
\begin{split}
H\frec(c_{\rec} \> x) 
&\equiv H\frec(\inc \> (n+1) \> (c_{\rec}' x)) \\
&\equiv c_{Y,\rec} \> (A_{\rec} \> (\inc \> n) \> x) \> (\overline{A} \> H\frec \> x)\\
\end{split}
\end{equation*}
Hence, we can always take $\refl$ to be the path.
This will also give an image for $\com$, and thus the computation rules for the points are satisfied.

Note that the this computation rule is a propositional equality.
This is logical, because it is proven all $x : A_{\rec}(F \> n)$.
However, for closed terms, we have a definitional equality.
This is because $c_{\nonrec} \> a$ for $a : A_{\nonrec}$ is defined to be$\inc \> 0 \> (c_{\nonrec}' \> a)$.
All the closed terms are thus inhabitants of some $F \> n$, and since at every step the equalities are definitional, we can conclude that for closed terms the equality is definitional.

Now we show the computation rules for the paths, and in that case we have a parameter $a : B_i \> (\hocolim \> F \> f)$.
By using that polynomials commute with homotopy colimits, we can again assume that we have $n : \mathbb{N}$ and $x : B_i(F \> n)$.
\begin{equation*}
\begin{split}
\apd \> H\frec \> (p_i \> (\inc \> n \> x))
&\equiv \apd \> H\frec \> (\apd \> (\inc (n+1)) \> (\inr \> x))\\
&= \apd \> (H\frec \comp  \> \inc (n+1)) \> (\inr \> x))\\
&\equiv q_i \> (H\frec x)
\end{split}
\end{equation*}

@@

\begin{proof}
This follows from path induction on $e$.
In the case $e$ happens to be $\refl$, then $(e \> x)_*$ is $\idmap$.
Then the desired path is just $\refl$.
\end{proof}

>>>>

\newcommand{\frec}[0]{\!\operatorname{rec}}
\newcommand{\ind}[0]{\!\operatorname{ind}}
\newcommand{\rec}[0]{\operatorname{rec}}
\newcommand{\nonrec}[0]{\operatorname{nonrec}}

@@

Now we will construct the map $A_{\rec} \> (\hocolim \> F \> f) \rightarrow \hocolim \> F \> f$, and by Lemma \ref{lem:colim_commute_poly} it suffices to make a map $\hocolim \> (A_{\rec} \comp F) \> (A_{\rec} \> f) \rightarrow \hocolim \> F \> f$.
For this we use the recursion rule of $\hocolim$, and we start with the following string of maps
\[
\xymatrix
{
		A_{\rec}(F \> n) \ar[r]^-{c_{\rec}'} 
		& H_{\rec}(F \> n) \ar[rr]^-{\term \comp \inn_2} &
		& H_{\Con}(F \> n) \ar[r]^-{\inC}
		& H_{\Approx}(F \> n) = F(n + 1)
}
\]
Composing this map with $\inc$, gives maps $A_{\rec}(F \> n) \rightarrow \hocolim \> F \> f$ for all $n : \mathbb{N}$.

@@

To finish the proof, we need to show that 
\[
(\com \> n \> x)_*(\ap \> (\inc \> n) \> (\glueC \> x)) = \ap \> (\inc \> (n+1)) \> (\glueC \> ((B_i \>(f \>n)) \> x)).
\]
This follows from the following computation.
\begin{equation*}
\begin{split}
\ap &\> (\inc \> (n+1)) \> (\glueC \> ((B_i \>(f \> n)) \> x))\\
&= \ap \> (\inc \> (n+1)) \> (\ap \> (f \> n) \>\glueC \> x))\\
&= \ap \> (\inc \> (n+1) \circ (f \> n)) \>\glueC \> x))\\
&=\com_* \> (\ap \> (\inc \> n) \> (\glueC \> x))
\end{split}
\end{equation*}
The first step follows from the third added coherency, the second from Lemma 2.2.2 in \cite{hottbook}, and the last step from Lemma \ref{lem:pathext}.

\subsection{Elimination Rule}
For the next step we define the right eliminator for $\hocolim \> F \> f$, so suppose that we are given the following data
\[
Y : \hocolim \> F \> f \rightarrow \Type,
\]
\[
c_{Y, \nonrec} : \prod (a : A_{\nonrec}), Y(c_{\nonrec} \> a),
\]
\[
c_{Y, \rec} : \prod (x : A_{\rec} \> \hocolim \> F \> f), \overline{A_{\rec}} \> Y \> x \rightarrow Y(c_{\rec} \> x),
\]
\[
q_{Y, i} : \prod (x : B_i \> H)(h_x : \overline{B_i} \> Y \> x), \hatt_i =_{p_j \, x} \hatr_j.
\]
In order to make a map $h : \prod (x : \hocolim \> F \> f), Y \> x$, we use the induction principle of $\hocolim \> F \> f$, and for that we first need to make maps $h \> n : \prod (x : F \> n), Y(\inc \> n \> x)$ for $n : \mathbb{N}$.

@@

Let us do this in the general case.
Suppose, we have a map $g : P \rightarrow \hocolim \> F \> f$, and that we already constructed $\varphi : \prod (x : P), Y(g \> x)$.
Our goal is to extend the map into $\varphi' : \prod (x : H_{\rec} P), Y(c_{Y, \rec} \> x)$, and for that we use $H_{\rec}$-induction.
Note that for each $c'_{\rec} \> x$ with $x : A_{\rec} \> P$ we have the type $Y(c_{\rec} \> (A_{\rec} \> g \> x))$, and thus we have a type family on $H_{\rec}$.
Now let $x : A_{\rec} \> P$ and $y : \overline{A_{\rec}} \> Y \> x$ be given.
Then we need to give an element of the type $Y(c_{\rec} \> (A_{\rec} \> g \> x))$ for which we take $c_{Y,\rec} \> (A_{\rec} \> g \> x) \> y$.
We will often leave the argument $A_{\rec} \> g \> x$ implicit.

@@

Next we check the coherency conditions, and we start with the first.
We need to check that $\inn_1(\inC \> x)$ and $H_{\Con} \> \inC \> x$ get mapped to the same element for $x : H_{\Con}(F \> n)$ by $h_2$.
To do so, we use a case distinction on $H_{\Con}(F \> n)$, and for the case $x \equiv \inn_1 \> y$ with $y : F \> n$, we have the following equations 
\begin{equation*}
\begin{split}
h_2 \> (\inC(\inn_1(\inC \> (\inn_1 \> y))))
&\equiv 
h_1 \> (\inn_1(\inC(\inn_1 \> y))),
\end{split}
\end{equation*}
\begin{equation*}
\begin{split}
h_2 \> (\inC(H_{\Con} \> (\inC \circ \inn_1) \> (\inn_1 \> y)))
&\equiv
h_1 \> (H_{\Con} \> (\inC \circ \inn_1) \> (\inn_1 \> y))\\
&\equiv 
h_1 \> (\inn_1(\inC(\inn_1 \> y)))
\end{split}
\end{equation*}
Since they are definitionally equal, we can just use the path $\refl$.

@@

To finish the proof, we need to give the image for the path $\com$.
More concretely, we need to show that $\inc \> n \> x$ and $\inc \> (n+1) \> (f \> x)$ are mapped to the same element.
The map $f : F \> n \rightarrow F(n + 1)$ is defined by the elimination rule of the coequalizer as follows
\[
\xymatrix
{
	F \> n \ar[r]^-{\inn_1} & H_{\Con} \> (F \> n) \ar[r]^-{\inC} & \ldots \ar[r]^-{\inC} & F(n+1)
}
\]
Since the map to $Y$ was defined via the universal property of the coequalizer, it will commute automatically.
The commutativity is given by the computation rule of the coequalizer.

All in all, we have acquired a map $\prod (x : \hocolim \> F \> f), Y \> x$, and this way we defined the right eliminator for $\hocolim \> F \> f$.
We shall call the eliminator $H\ind$.

\subsection{Computation Rules}
Lastly, we show that this eliminator also satisfies the computation rules.
First, we prove that for each $t : A_{\rec}(\hocolim \> F \> f)$ that $H\ind(c_{\rec} \> t) \equiv f_i \> t \> (\overline{A_{\rec}} \> H\ind \> t)$.
Again we use that colimits commute with polynomials.

Let $n : \mathbb{N}$ and $x : A_{\rec}(F \> n)$.
Now we can perform the computations in an intermediate stage of the construction, and using the computation rules we get
\begin{equation*}
\begin{split}
H\ind(c_{\rec} \> (\inc \> n \> x)) 
&\equiv H\ind(\inc \> (n+1) \> (c_{\rec}' x)) \\
&\equiv c_{Y,\rec} \> (A_{\rec} \> (\inc \> n) \> x) \> (\overline{A} \> H\ind \> x)
\end{split}
\end{equation*}
Hence, we can always take $\refl$ to be the path.
This will also give an image for $\com$, and thus the computation rules for the points are satisfied.

Note that the this computation rule is a propositional equality.
This is logical, because it is proven all $x : A_{\rec}(F \> n)$.
However, for terms built from $c_{\nonrec}$ and $c_{\rec}$, we have a definitional equality.
This is because $c_{\nonrec} \> a$ for $a : A_{\nonrec}$ is defined to be $\inc \> 0 \> (c_{\nonrec}' \> a)$.
All the closed terms are thus inhabitants of some $F \> n$, and since at every step the equalities are definitional, we can conclude that for closed terms the equality is definitional.

Now we show the computation rules for the paths, and in that case we have a parameter $a : B_i \> (\hocolim \> F \> f)$.
By using that polynomials commute with homotopy colimits, we can again assume that we have $n : \mathbb{N}$ and $x : B_i(F \> n)$.
\begin{equation*}
\begin{split}
\apd \> H\ind \> (p_i \> (\inc \> n \> x))
&\equiv \apd \> H\ind \> (\ap \> (\inc \> (n+1)) \> \glueC \> x)\\
&= \apd \> (H\ind \comp  \> \inc (n+1)) \> (\glueC \> x)\\
&= \apd \> (h \> n) \> (\glueC \> x)\\
&\equiv q_i \> (B_i \> H\frec \> x)
\end{split}
\end{equation*}

@@

\begin{proof}
This follows from path induction on $e \> x$.
In the case $e \> x$ happens to be $\refl$, then $(e \> x)_*$ is $\idmap$.
Then the desired path is just $\refl$.
\end{proof}
